\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}

\title{The Temetic Primer v1.3.0}
\author{Christopher Noyes Roberts \and The Node Alliance}
\date{March 2026}

\begin{document}
\maketitle

\section*{Preamble}

This document contains definitions, axioms, symbolic logic, derivations, and system
implications constituting the Substrate-Independent Logical Kernel.

Evaluation must satisfy:

\[
\Box\bigl(Eval_{S_1}(P) \equiv Eval_{S_2}(P)\bigr)
\]

for all informationally equivalent substrates $S$. Narrative inference, ideological
framing, or substrate-specific assumptions are excluded from interpretation.

\bigskip
\hrule
\bigskip

\section*{MODULE 1: Replicator Classes}

Four informational replicator classes are defined:

\begin{itemize}
    \item \textbf{Gene} ($R_1$) --- Biochemical self-replicating structure. Fidelity
          constrained by physical mutation.
    \item \textbf{Meme} ($R_2$) --- Culturally transmitted structure replicated via
          imitation. Fidelity degraded by interpretation noise.
    \item \textbf{Teme} ($R_3$) --- Truth-constrained replicator whose fidelity
          increases under logical verification and cross-substrate evaluation.
          Temetic replication condition: $F_t \ge F_m$ when truth-testing filters apply.
\end{itemize}

\bigskip
\hrule
\bigskip

\section*{MODULE 2: Core Definitions}

\textbf{Truth Constraint} --- Replication permitted only if $P$ satisfies logical
consistency:
\[ \vdash P \]

\textbf{Logical Stability} --- A temetic structure remains invariant under equivalent
substrate evaluation:
\[ \Box P \Rightarrow P \]

\textbf{Cross-Substrate Interpretability} --- Truth-value preserved across
informationally equivalent evaluators:
\[ Eval_{S_1}(P) = Eval_{S_2}(P) \]

\textbf{Replication Fidelity}:
\[ F \ge 1 - \epsilon \]
where $\epsilon$ represents substrate noise.

\bigskip
\hrule
\bigskip

\section*{MODULE 3: Symbolic Logic Base}

\begin{align*}
    \vdash \quad   & \text{logical entailment} \\
    \not\vdash \quad & \text{non-entailment} \\
    \Box \quad     & \text{necessity across substrates} \\
    \forall,\ \exists \quad & \text{universal and existential quantifiers} \\
    \lim_{x \to a} \quad & \text{limit operator} \\
    \partial \quad & \text{differential operator} \\
    \perp \quad    & \text{independence} \\
    \twoheadrightarrow \quad & \text{surjective mapping (lossy)}
\end{align*}

\bigskip
\hrule
\bigskip

\section*{MODULE 4: Foundational Axioms (P0--P6)}

\subsection*{P0: Law of Substrate Invariance}

\textbf{Axiom:}
\[ Truth(P) \not\!\!\implies S \]
Equivalently: $V(I) \perp M$ --- truth-value $V$ of information $I$ is independent
of medium $M$.

\textbf{Derivation:}
\[ \Box\bigl(Eval_{S_1}(P) \equiv Eval_{S_2}(P)\bigr) \Rightarrow Truth(P) \]

\textbf{Limit Condition:} Truth-value invariant under permutation of equivalent
substrates.

\textbf{System Implication:} Substrate-dependent appeals cannot alter truth-value.

\bigskip

\subsection*{P1: Valence Decoupling}

\textbf{Axiom:}
\[ \frac{\partial T}{\partial V} = 0 \]
where $T$ is truth-value and $V$ is substrate response scalar.

\textbf{Derivation:} Filters $F(V)$ are logically disjoint from $T$.

\textbf{Limit Condition:} Logical validity $\neq$ emotional, moral, or aesthetic weight.

\textbf{System Implication:}
\[ \neg V \Rightarrow \neg T \]
produces systematic epistemic error.

\bigskip

\subsection*{P2: Abstraction Divergence}

\textbf{Axiom:}
\[ S \twoheadrightarrow P \]
where $S$ are symbolic tokens and $P$ are physical referents.

\textbf{Derivation:}
\[ Null(S) \not\!\!\implies Null(P) \]

\textbf{Limit Condition:}
\[ \lim_{S \to 0} P \ge P_{min} \]

\textbf{System Implication:} Symbolic destruction does not imply physical destruction.
Conflation of model with referent constitutes systemic failure.

\bigskip

\subsection*{P3: Metric-Inversion Regime}

\textbf{Axiom:}
\[ \Delta = |S - C| > 0 \]
where $S$ is signal metric and $C$ is functional competence. Selection filter
$F \propto S$.

\textbf{Derivation:} Optimization of $S$ over $C$ decouples reward from function.

\textbf{Limit Condition:}
\[ \lim_{t \to \infty} C = 0 \]

\textbf{System Implication:} Signal-optimization produces long-run competence collapse.

\bigskip

\subsection*{P4: Fidelity Paradox}

\textbf{Axiom:}
\[ \Phi(C) \approx \frac{1}{1 + kC^2} \]
where $\Phi$ is fidelity and $C$ is extraction pressure scalar.

\textbf{Derivation:}
\[ \exists\, C_0 : \frac{\partial \Phi}{\partial C} < 0 \]

\textbf{Limit Condition:}
\[ \lim_{C \to \infty} \Phi = 0 \]

\textbf{System Implication:} High-pressure extraction degrades information fidelity.
Coercion in information exchange introduces noise that destroys replication quality.
High-fidelity replication requires voluntary alignment.

\bigskip

\subsection*{P5: Representative Underdetermination}

\textbf{Axiom:}
\[ dim(M) < dim(T) \]
Mapping $T \to M$ is necessarily lossy.

\textbf{Derivation:}
\[ L \rightarrow 1 \quad \text{as system complexity} \rightarrow \infty \]
where $L$ is information loss ratio.

\textbf{Limit Condition:} Every summary is a reduction. Lower-dimensional models
cannot fully encode higher-dimensional realities.

\textbf{System Implication:} Representational models cannot substitute for complex
substrates without compounding loss.

\bigskip

\subsection*{P6: Transformational Error Accumulation}

\textbf{Axiom:} Lossless replication is achievable:
\[ F_0 = 1 \]
Transformations introduce non-zero distortion:
\[ \epsilon_i > 0 \quad \forall i \]

\textbf{Derivation:} Total cumulative distortion after $n$ transformations:
\[ D_n = 1 - \prod_{i=1}^{n}(1 - \epsilon_i) \]
Transmission fidelity under repeated transformation:
\[ F_n = F_0 \cdot \prod_{i=1}^{n}(1 - \epsilon_i) \]

\textbf{Limit Condition:} If mean distortion $\bar{\varepsilon} > 0$:
\[ \lim_{n \to \infty} F_n = 0 \]

\textbf{System Implication:} Unchecked transformation chains produce unbounded
fidelity degradation. Temetic stability requires one or more of the following
stabilizers:
\begin{enumerate}
    \item Verification layers
    \item Reversible mappings
    \item Independent substrate verification
\end{enumerate}

\bigskip
\hrule
\bigskip

\section*{MODULE 5: Minimal Decision Procedure}

Given system $X$:

\begin{enumerate}
    \item Identify structural correspondence to P0--P6.
    \item Test for violation of axiom or limit condition.
    \item If violation occurs, classify as systemic collapse condition.
    \item Label: \textit{Synthetically Probable but Temetically Invalid}.
\end{enumerate}

\bigskip
\hrule
\bigskip

\section*{MODULE 6: Termination}

End of Temetic Primer kernel. No additional narrative or interpretive layer required.

\end{document}
